Cryptobenthic reef assemblages are more distinct across a 90 m depth gradient than 2,500 km of shallow marine habitat in the Hawaiian Archipelago

### 2024-09-18 manuscript.Rmd
some updates

A total of 89 ARMS units were deployed over the course of six years (2009--2016) in shallow coral reef habitat (8--17 m depth) in both the Main (MHI) and Northwest Hawaiian Islands (NWHI) ([@fig:fig_map; @tbl:arms_table]). Only units deployed on outer reefs were included in this study. To explore the effects of depth, we also present data from an earlier study by Hurley et al. [-@hurleyAssessmentShallowMesophotic2016], in which 27 units were deployed off O&#x02BB;ahu from 2009--2012 in 12--90 m depth; nine units were deployed at 12 m, and six each at 30 m, 60 m, and 90 m ([@tbl:arms_table]). Shallow units were deployed according to NOAA-RAMP methodology, whereas units $\geq$ 30 m were deployed and recovered using technical mixed gas closed-circuit rebreather diving [see @hurleyAssessmentShallowMesophotic2016].

```{r fig-map, fig.cap='Map of ARMS deployment sites across the Hawaiian Archipelago (Papahānaumokuākea Marine National Monument (NWHI) is shaded in light blue).', results="asis", fig.width=9, fig.height=6, dependson="init"}
library(sf)
library(rnaturalearth)

@@

ARMS units were retrieved by encapsulating them in-situ in a 100 μm mesh-lined crate to prevent escape of motile organisms. Once returned to the surface, they were placed in containers of aerated, filtered seawater, crates were removed, and units were systematically disassembled plate-by-plate [see @lerayDNABarcodingMetabarcoding2015 for processing details]. All adult brachyuran crabs ($\geq$ 2 mm) were photographed, identified, and preserved in 95% ethanol. Specimens were initially identified to family or genus using Poore & Ahyong [-@pooreMarineDecapodCrustacea2023] and additionally to species where possible according to expert advice from Scott Godwin (former curator of invertebrate zoology at Bishop Museum) and Gustav Paulay (curator of invertebrate zoology, Florida Museum of Natural History). Where species-level identifications were not possible, taxa were collapsed to family or genus (depending on individual identification). We recognize that the species-level taxonomy of Indo-Pacific brachyurans remains poorly resolved, particularly for abundant groups like the Chlorodiellinae [@lasleyjrIntegrativeSystematicRevision2014; @lasleyjrPhylogeneticRelationshipsUbiquitous2015], and some "species" are known to comprise complexes [e.g. *Perinia tumida*, @servisCharacterizingCoralReef2020; @plaisanceReefassociatedCrustaceanFauna2009]. However, the prevalence of complexes and cryptic diversity suggests that our assignments are likely to under- rather than overrepresent the true number of taxa. Thus, any significant patterns detected likely persist in spite rather than because of cryptic diversity.

@@

Unless otherwise specified, all analyses were performed using `R` version 4.1.2 [@rcoreteamLanguageEnvironmentStatistical2021]. Community matrices were loaded and encapsulated into data objects using the R package `phyloseq` [@mcmurdiePhyloseqPackageReproducible2013] and multivariate community analyses were performed using the package `vegan` [@oksanenVeganCommunityEcology2022] on Bray-Curtis dissimilarities of Hellinger-transformed species abundance matrices [@legendreEcologicallyMeaningfulTransformations2001; @raoReviewCanonicalCoordinates1995].

@@

Potential drivers of brachyuran crab community structure were assembled from previous studies [@selkoeEvaluatingAnthropogenicThreats2008; @halpernGlobalMapHuman2008; @selkoeDNACoralReef2016; @selkoeMapHumanImpacts2009], NOAA's National Coral Reef Monitoring Program (NCRMP), the Pacific Islands Ocean Observing System (PacIOOS), and MODIS satellite data products compiled by NASA's Physical Oceanography/Ocean Biology Distributed Active Archive Centers [@nasagoddardspaceflightcenterModerateresolutionImagingSpectroradiometer2022; @nasaobpgMODISAquaGlobal2020]. Variables were either intrinsic (site depth, latitude, longitude, island slope) or derived from remote sensing and/or modeling (chlorophyll-A, sea-surface temperature, coral cover, potential larval immigration, cumulative human impacts) ([@tbl:env_table; @tbl:arms_metadata_table]). Sea-surface temperature (SST) and chlorophyll-A values were calculated by averaging monthly means across the duration of the soaking period at each sampling site using AQUA MODIS 4 km-resolution gridded datasets. The cumulative human impacts index, a dimensionless numerical value representing spatially-explicit cumulative anthropogenic impacts on various marine ecosystems, is based on a synthesis of a combination of "high resolution, annual data on the intensity of 14 human stressors and their impact on 21 marine ecosystems over 11 years (2003--2013)" [@halpernGlobalMapHuman2008; @halpernRecentPaceChange2019]. For details regarding data import, preparation, and synthesis, see Supplemental material. Due to logistical limitations, environmental data were collected only for shallow samples (\< 30 m).

@@

For the complete dataset, there was no significant difference in observed species richness between the MHI and NWHI ([@fig:fig_alpha[]{}a]; `r k_test(all_alpha$k_r_island)`), but effective Simpson diversity was significantly different ([@fig:fig_alpha[]{}c]; `r k_test(all_alpha$k_s_island,plain=TRUE)`). Richness was not significantly different between shallow and deep samples ([@fig:fig_alpha[]{}b]; `r k_test(all_alpha$k_r_depth,plain=TRUE)`) but effective Simpson diversity was ([@fig:fig_alpha[]{}d]; `r k_test(all_alpha$k_s_depth,plain=TRUE)`). When only shallow arms \< 30 m were considered, there were no significant differences in alpha diversity between the MHI and NWHI for either index (richness: `r k_test(shallow_alpha$k_r_island,plain=TRUE)`, [@fig:fig_alpha2[]{}a]; effective Simpson: `r k_test(shallow_alpha$k_s_island,plain=TRUE)`, [@fig:fig_alpha2[]{}b]).

@@

<div class="multi-refs"></div>

@@

## Processing of satellite-derived environmental data
Level-3 mapped (4 km grid-size) sea-surface temperature (SST) and chlorophyll-A (chl-A) data were retrieved from NASA's Physical Oceanography Distributed Active Archive Center [PO.DAAC; @nasaobpgMODISAquaGlobal2020] and Ocean Biology Distributed Active Archive Center [OB.DAAC; @nasagoddardspaceflightcenterModerateresolutionImagingSpectroradiometer2022] respectively. Monthly summary datasets were downloaded for the period of 2009--2016 and averaged over each of the ARMS unit soaking periods (e.g. 2009--2011, 2010--2012, etc.). Values (SST or chl-A) were extracted for the geographic coordinates of each sampling site from the averaged dataset for the relevant soaking period. Specific monthly-average dataset names can be found in @tbl:nasa_table.

## Processing of cumulative human impacts data
Gridded cumulative human impacts raster data calcuated by Halpern et al [-@halpernRecentPaceChange2019] was downloaded from the Knowledge Network for Biocomplexity (KNB) repository [@frazierRecentPaceChange2019]. Human impacts datasets were downloaded for each available year in this study (2009--2013) and averaged over ARMS unit soaking periods (as above). Since cumulative impacts data was only available up until 2013, we used the combined average of all available years for the 2013--2016 soaking period. Since the availability of data points at individual coordinates was highly variable, we averaged impact scores for each sampled island in the study.

@@

```{r fig-alpha2, fig.cap='Box plots of (a) species richness and (b) Effective Simpson diversity by island group for shallow samples only. Notches as in [@fig:fig_alpha]. Both comparisons are non-significant.', results="asis", dev="svg", fig.width=8, fig.height=4, dependson="init"}
art <- shallow_alpha$richness_table %>% 
  mutate(island_group = str_to_sentence(island_group)) %>%
  mutate(shallow_deep = str_to_sentence(shallow_deep),shallow_deep = fct_reorder(shallow_deep,depth))

@@

<div class="multi-refs"></div>

>>>>

A total of 89 ARMS units were deployed over the course of six years (2009--2016) in shallow coral reef habitat (`r format_range(cc$crabs_shallow %>% sample_tibble() %>% pull(depth))` m depth) in both the Main (MHI) and Northwest Hawaiian Islands (NWHI) ([@fig:fig_map; @tbl:arms_table]). Only units deployed on outer reefs were included in this study. To explore the effects of depth, we also present data from an earlier study by Hurley et al. [-@hurleyAssessmentShallowMesophotic2016], in which 27 units were deployed off O&#x02BB;ahu from 2009--2012 in `r format_range(cc$crab_data %>% sample_tibble() %>% filter(island == "Oahu") %>% pull(depth))` m depth; nine units were deployed at 12 m, and six each at 30 m, 60 m, and 90 m ([@tbl:arms_table]). Shallow units were deployed according to NOAA-RAMP methodology, whereas units $\geq$ 30 m were deployed and recovered using technical mixed gas closed-circuit rebreather diving [see @hurleyAssessmentShallowMesophotic2016]. Unless referring to specific depth zones, we will refer to any units deployed \< 30 m as "shallow" and those deployed $\geq$ 30 m as "deep".

```{r fig-map, fig.cap='Map of ARMS deployment sites across the Hawaiian Archipelago (Papahānaumokuākea Marine National Monument (NWHI) is shaded in light blue).', results="asis", fig.width=9, fig.height=6, dependson="init"}
library(sf)
library(rnaturalearth)
library(rnaturalearthhires)

@@

ARMS units were retrieved by encapsulating them in-situ in a 100 μm mesh-lined crate to prevent escape of motile organisms. Once returned to the surface, they were placed in containers of aerated, filtered seawater, crates were removed, and units were systematically disassembled plate-by-plate [see @lerayDNABarcodingMetabarcoding2015 for processing details]. All adult brachyuran crabs ($\geq$ 2 mm) were photographed, identified, and preserved in 95% ethanol. Specimens were initially identified to family or genus using Poore & Ahyong [-@pooreMarineDecapodCrustacea2023] and additionally to species where possible according to expert advice from Scott Godwin (former curator of invertebrate zoology, Bishop Museum) and Gustav Paulay (curator of invertebrate zoology, Florida Museum of Natural History). Where species-level identifications were not possible, taxa were collapsed to family or genus (depending on individual identification). We recognize that the species-level taxonomy of Indo-Pacific brachyurans remains poorly resolved, particularly for abundant groups like the Chlorodiellinae [@lasleyjrIntegrativeSystematicRevision2014; @lasleyjrPhylogeneticRelationshipsUbiquitous2015], and some "species" are known to comprise complexes [e.g. *Perinia tumida*, @servisCharacterizingCoralReef2020; @plaisanceReefassociatedCrustaceanFauna2009]. However, the prevalence of complexes and cryptic diversity suggests that our assignments are likely to under- rather than overrepresent the true number of taxa. Thus, any significant patterns detected likely persist in spite of rather than because of cryptic diversity.

@@

Unless otherwise specified, all analyses were performed using `R` version `r R.version$major`.`r R.version$minor` [@rcoreteamLanguageEnvironmentStatistical2021]. Community matrices were loaded and encapsulated into data objects using the R package `phyloseq` [@mcmurdiePhyloseqPackageReproducible2013] and multivariate community analyses were performed using the package `vegan` [@oksanenVeganCommunityEcology2022] on Bray-Curtis dissimilarities of Hellinger-transformed species abundance matrices [@legendreEcologicallyMeaningfulTransformations2001; @raoReviewCanonicalCoordinates1995].

@@

Potential drivers of brachyuran crab community structure were assembled from previous studies [@selkoeDNACoralReef2016; @halpernRecentPaceChange2019], NOAA's National Coral Reef Monitoring Program (NCRMP), the Pacific Islands Ocean Observing System (PacIOOS), and MODIS satellite data products compiled by NASA's Physical Oceanography/Ocean Biology Distributed Active Archive Centers [@nasagoddardspaceflightcenterModerateresolutionImagingSpectroradiometer2022; @nasaobpgMODISAquaGlobal2020]. Variables were either intrinsic (site depth, latitude, longitude, island slope) or derived from remote sensing and/or modeling (chlorophyll-A, sea-surface temperature, coral cover, potential larval immigration, cumulative human impacts) ([@tbl:env_table; @tbl:arms_metadata_table]). Sea-surface temperature (SST) and chlorophyll-A values were calculated by averaging monthly means across the duration of the soaking period for each deployment using AQUA MODIS 4 km-resolution gridded datasets. The cumulative human impacts index, a dimensionless numerical value representing spatially-explicit cumulative anthropogenic impacts on various marine ecosystems, is based on a synthesis of a combination of "high resolution, annual data on the intensity of 14 human stressors and their impact on 21 marine ecosystems over 11 years (2003--2013)" [@halpernGlobalMapHuman2008; @halpernRecentPaceChange2019]. The index of potential larval immigration was imported from [@selkoeDNACoralReef2016], and represents an "estimate of in-closeness centrality" from a biophysical model simulating larval dispersal in the Hawaiian Islands [@selkoeDNACoralReef2016 supp. mat.] For details regarding data import, preparation, and synthesis, see Supplemental material. Due to logistical limitations, environmental data were collected only for shallow samples (\< 30 m).

@@

For the complete dataset, there was no significant difference in observed species richness (`r k_test(all_alpha$k_r_island)`; [@fig:fig_alpha[]{}a]) or effective Simpson diversity (`r k_test(all_alpha$k_s_island,plain=TRUE)`; [@fig:fig_alpha[]{}c]) between the MHI and NWHI. Richness was not significantly different between shallow and deep samples (`r k_test(all_alpha$k_r_depth,plain=TRUE)`; [@fig:fig_alpha[]{}b]) but effective Simpson diversity was (`r k_test(all_alpha$k_s_depth,plain=TRUE)`; [@fig:fig_alpha[]{}d]). When only shallow arms \< 30 m were considered, there were no significant differences in alpha diversity between the MHI and NWHI for either index (richness: `r k_test(shallow_alpha$k_r_island,plain=TRUE)`, [@fig:fig_alpha2[]{}a]; effective Simpson: `r k_test(shallow_alpha$k_s_island,plain=TRUE)`, [@fig:fig_alpha2[]{}b]).

@@

<!-- show main bibliography -->
<div class="multi-refs"></div>

@@

## Data import and processing

All data import and processing code for environmental variables used in the dbRDA model can be found in the file `get-data.R` from the associated github repository. Details for specific datasets are given below.

### Processing of satellite-derived environmental data
Level-3 mapped (4 km grid-size) sea-surface temperature (SST) and chlorophyll-A (chl-A) data were retrieved from NASA's Physical Oceanography Distributed Active Archive Center [PO.DAAC; @nasaobpgMODISAquaGlobal2020] and Ocean Biology Distributed Active Archive Center [OB.DAAC; @nasagoddardspaceflightcenterModerateresolutionImagingSpectroradiometer2022] respectively. Monthly summary datasets were downloaded for the period of 2009--2016 and averaged over each of the ARMS unit soaking periods (e.g. 2009--2011, 2010--2012, etc.). Values (SST or chl-A) were extracted for the geographic coordinates of each sampling site from the averaged dataset for the relevant soaking period. Original raw monthly-average dataset names can be found in @tbl:nasa_table.

### Processing of cumulative human impacts data
Gridded cumulative human impacts raster data calcuated by Halpern et al [-@halpernRecentPaceChange2019] was downloaded from the Knowledge Network for Biocomplexity (KNB) repository [@frazierRecentPaceChange2019]. Human impacts datasets were downloaded for each available year in this study (2009--2013) and averaged over ARMS unit soaking periods (as above). Since cumulative impacts data was only available through 2013, we used the combined average of all available years for the 2013--2016 soaking period. Since the availability of data points at individual coordinates was highly variable, we averaged impact scores for each sampled island in the study.

### Other datasets
Coral cover and potential larval immigration were imported from the online supplemtal material accompanying Selkoe et al. [-@selkoeDNACoralReef2016].

## Figures and tables

@@

```{r fig-alpha2, fig.cap='Box plots of (a) species richness and (b) Effective Simpson diversity by island group for shallow samples only. Notches as in [@fig:fig_alpha]. Both comparisons were non-significant.', results="asis", dev="svg", fig.width=8, fig.height=4, dependson="init"}
art <- shallow_alpha$richness_table %>% 
  mutate(island_group = str_to_sentence(island_group)) %>%
  mutate(shallow_deep = str_to_sentence(shallow_deep),shallow_deep = fct_reorder(shallow_deep,depth))

@@

<!-- show supplemental bibliography -->
<div class="multi-refs"></div>
